---
title: "Week 07 Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Exercises

```{r}
if(! require("pacman")) install.packages("pacman")

pacman::p_load(
  tidyverse,
  here
)

source(here::here("hw_functions.R"))
```

1.  Describe some of the benefits of using functions.

Functions can be used to make code more compact, readable, and reproducible, especially when an operation is repeated. Generalizing a particular operation with a function means ability to consistently repeat steps without typos/ordering errors for calculations, data cleaning, and data processing.


2.  Practice turning the following code snippets into functions. Think about what each function does. What would you call it? How many arguments does it need? Can you rewrite it to be more expressive or less duplicative?

```{r}
mean(is.na(x))

x / sum(x, na.rm = TRUE)

sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)
```

Each of these only needs one argument in the form of a vector. The first and third operations produce single summary numbers, while the second operation outputs a vector.
```{r}
# fraction of entries with contents NA
fraction_na <- function(field){
  mean(is.na(field))
}

# fraction of the column total for each entry
fraction_of_total <- function(field){
  field / sum(field, na.rm = TRUE)
}

# ratio of column standard deviation to the column mean
sd_mean_ratio <- function(field){
  sd(field, na.rm = TRUE) / mean(field, na.rm = TRUE)
}

#Test data generation
sample(1:100, 25, replace = T)  # generate radome 25 mumbers

test_data  <- c(50, 31, 69, 22, 30, 72, 63, 44, 13, 31, 16, 50, NA, 56, 72, 65, 16, NA, 10, 93, 38, 14, 91, 53, 82, 91, 81)   # Copy 25-member list and insert 2 NAs; 27 total entries
```
Now that we defined functions, let's test...
```{r}
fraction_na(test_data)  # function
2/27                    # expected value from test data

fraction_of_total(test_data)

sd_mean_ratio(test_data)

```
`fraction_na` produces the expected result from the input data

`fraction_of_total` produces a vector with each entry's fraction of the total 

`sd_mean_ratio` produces a single value between 0 and 1, as expected. Based on the range of data, a standard deviation of approx half the mean is reasonable.

The only repeated functionality within these operations is the `na.rm` argument; it's not clear to me whether the argument could be turned into a function or that doing so could generate more compact code in this case.



3.  Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
f3 <- function(x, y) {
  rep(y, length.out = length(x))
}
```
f1 --> repl_prefix
f1 replaces the beginning of an input string with a specified prefix

f2 --> drop_last
f2 returns the vector without the last element, including a NULL where the length of the initial vector is 1.

f3 --> trunc_fill
f3 returns a vector in which the length of vector y is truncated or content of y is repeated to match the length of vector x.

4.  Implement a fizzbuzz function. It takes a single number as input. If the number is divisible by three, it returns "fizz". If it's divisible by five it returns "buzz". If it's divisible by three and five, it returns "fizzbuzz". Otherwise, it returns the number. Make sure you first write working code before you create the function.

```{r}
# establish function

fizzbuzz <- function(num){
      # is.integer() does not detect integer results
      # is.wholenumber code from CRAN help file on Integer Vectors
  is.wholenumber <-
      function(x, tol = .Machine$double.eps^0.5)  abs(x - round(x)) < tol

  #recursive if-else comparison
  if(is.wholenumber(num/15)) {    # divisible by both 3 & 5 = divvisible by 15
      print("fizzbuzz")
  } else {
    
    if(is.wholenumber(num/5)) { # order of 3 & 5 comparison unimportant
          print("buzz")
      } else {

        if(is.wholenumber(num/3)) {
            print("fizz")
          } else {
            print(num)          # when all comparisons fail, print num
          }
       }
  }
}
```
```{r}
# test function
fizzbuzz(30)
fizzbuzz(18)
fizzbuzz(25)
fizzbuzz(16)
fizzbuzz(35.0000001)  # Decimal limit of is.wholenumber() with provided tolerance function; if more precision is needed, adjust tolerance. Note R only prints 5 decimal digits, and omits these if the decimal digits are all 0.
```


We'll be working with national baby names data from the [Social Security Administration](https://www.ssa.gov/oact/babynames/limits.html)

5. The following code will create your data frame:

```{r US names}
files <- list.files(here::here("Data"), pattern = "\\.txt", full.names = T)

names_df <- map_df(files, ~read_name(.))
```

Describe what the `read_name` function is doing.

`read_name` as given in hw_functions.R is doing the following:
+ extracting year from the data file name by a 2-step process
+ creating a data frame by reading data from text files in specified path as CSV, assigning column headings
+ changing the year from a 4-digit number to a date corresponding to Jan 1 of that year
+ creating a rank column for each year based on the counts for ech gender




6.  Create a function that allows a user to select a name (or names) and graph its trend over time.

* If multiple names are selected, they should each be represented on the graph as a unique color.
* The user should be able to select whether they want to graph based upon the `count` or the `rank` of the name.
* The user should be able to indicate if they are interested in male names, female names, or either.

```{r}
glimpse(names_df)

# working example to build function
names_df %>%
  filter(name == "Mary"| name == "Ann") %>%   # this syntax proved important
  filter(gender == "F" | gender == "M")  %>%  # vs using a c() style list
  
  group_by(year, name) %>%
  mutate(count_total = sum(count)) %>%      # sum for M and F to avoid janky
  ungroup() %>%
  
  ggplot(aes(x = year,
             y = count_total,
             color = name,
             group = name)) +
  geom_line() +
  #scale_y_log10(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  theme(legend.position = "top",
        axis.title = element_text(face = "bold"),
        axis.text =  element_text(face = "italic"))  +
  labs(color = "")
```

```{r}
# sandbox
ycolgen <- function(df, popularity){
        if(popularity == "count")
        mutate(df, yplotcol = sum(count))
        else
        mutate(df, yplotcol = max(rank))  
      }

names_df %>% 
ycolgen(popularity = "count")


```

```{r}
#build function
namesplot<- function(df, names, popularity = "count", single_gender = NULL){
  
# set function to filter gender
    gender_filter <- function(df, single_gender){
      if(is.null(single_gender)){
      filter(df, gender == "F" | gender == "M")
      } else {
      filter(df, gender == single_gender) }
    }
    
# setup sub-function to mutate ycol
    # generate y col to plot
    # sum or rankmax for M and F to avoid janky effect of 2 points per year
    # calc has no effect on single-sex data

     ycolgen <- function(df, popularity){
        if(popularity == "count")
        mutate(df, yplotcol = sum(count))
        else
        mutate(df, yplotcol = max(rank))  
      }  
  
   #load data & start filtering
    df %>%  
      filter(name == {{names}})  %>%
      gender_filter(single_gender)  %>% 


    # generate y column
      group_by(year, name) %>%
      ycolgen({{popularity}})%>%
      ungroup()  %>%
  
    
  ggplot(aes(x = year,
             y = yplotcol,
             color = name,
             group = name)) +
  geom_line()  +
  
  scale_y_continuous(labels = scales::comma) + 

  theme_minimal()  +
  theme(legend.position = "top",
        axis.title = element_text(face = "bold"),
        axis.text =  element_text(face = "italic"))  +
  labs(x = "",
       y = popularity,
       color = "")
  
  }

```
```{r}
# test function -- one name
names_df %>%
namesplot("Ann", "rank", "F")
```
Requesting one name with all explicit arguments works!
```{r}
# test function -- one name, implicit arguments
names_df %>%
namesplot("Ann")   
```
Requesting one name with remaining arguments implicit works!
```{r}
# test function -- one name, implicit arguments
names_df %>%
namesplot("Howard")   
```
```{r}
# test function -- more names
names_df %>%
namesplot(c("Ann","Mary"), "count", "F")   
```
Two names with all explicit arguments throws an error related to the names but appears to work.
```{r}
# test function -- more names
names_df %>%
namesplot(c("Ann","Mary"), "count")   
```
Combining multiple names with gender = NULL causes problems. Including "count" or leaving it as a blank default has no effect. 
```{r}
# test function -- more names
names_df %>%
namesplot(c("Ann","Mary"), "count", c("M","F"))   
```
```{r}
# test function -- more names
names_df %>%
namesplot(c("Ann","Howard"), "count", c("M","F"))   
```
Explicitly saying we're looking for both M and F causes similar problems that seem to have a different shape.

I think this goes back to the syntax `filter(field == "target1" | field == "target2" | ...)` that works best. It looks like it is variably generating OR or AND on these arguments (and maybe a stray NOR in there...) instead of looking for all instances of both items in the list. The NULL gender input was reasonable to code in ; however, unfortunately I was not able to track down how to generate this syntax within the function from the inputs. That's what I would try next; my guess is that it's probably an iterative regex lookup.



7. Create a repository for your course project and share it with @rharrington31. Your project should include:

* A `README.md` file
* A script (whether it is a script or a markdown file) that successfully ingests your data for your analysis.

The repository is at https://github.com/d-conner/NEH_grants